\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-08-11 paper/paper.tex
Reworking the explanation of model assessment section

\subsection{Results}
\subsubsection{Model Assessment}
To assess the goodness of fit for our final model (Model 4), we use a graphical posterior predictive check. As stated by \citet{gelman1996postpred}, a graphical comparison of replicated data sets from the fitted model with the observed data can be very informative with regard to identification of problems with the model.

To generate replicate data sets, we must first decide what to do with regard to censoring and truncation. The representation we choose to display the data, non-parametric probability plots, guides us in making this decision. The variability of this estimator arises primarily from the number of ``at risk" units at any given time, which depends in part on these processes. Since we are not modeling the mechanisms of truncation nor censoring, we want to try to duplicate these mechanisms in the replicated data. Our method for generating a replicate data set  $y_{rep,g,i}^{(s)}$ for $s = 1 \dots 19$ is as follows:
\begin{enumerate}
\item Simulate $\tilde{y}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$.  Where each set of parameters is randomly drawn from the full posterior distribution.
\item If $\tilde{y}_{rep,g,i} < t_{g,i}^L$, then remove this time as this record is not observed
\item Define $m_{g,i}>0$ for each $g$ and $i$. If $c_{g,i}=0$, let $m_{g,i}=\max \{y_{G,i}: G=g\}$, else let $m_{g,i}=y_{g,i}$.\\
If $\tilde{y}_{rep,g,i}^{(s)}>m_{g,i}$, set $y_{rep,g,i}^{(s)}=m_{g,i},\; c_{rep,g,i}^{(s)}=1$, otherwise set $y_{rep,g,i}^{(s)}=\tilde{y}_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=0$.
\end{enumerate}

We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to reestimate the Kaplan-Meier nonparametric estimator. The resulting plots are displayed in figure \ref{fig:post-pred-KM}. For most drive-models, the observed estimates appear to behave similarly to those derived from the replicated data. As discussed in \ref{subsec:ex1}, drive-model 14 has a discrepancy in the right tail, which we addressed there. Drive-model 4 also appears as an exception, which we discuss now.

@@

To better understand the reason for the discrepancy between the observed data for drive-model 4 and its posterior predictive distribution, we plot the raw data along with the truncation-adjusted nonparametric estimate of time to failure and the posterior estimate (see figure \ref{fig:ex-mod-4}). From these plots, it is clear that the large discrepancy between the posterior and the nonparametric estimate is driven by a small number of failures that occur in the first quarter year. Indeed, if we ignore records with observation periods that include the first quarter, the modified nonparametric estimate falls within the posterior 90\% pointwise credible band. Our trouble with fitting this drive-model is that while there is evidence of infant mortality in the relatively few units that were observed in the first quarter, the problem seems to go away much quicker than would be expected if there were a large number of units susceptible to the early failure mode. In any case, this example demonstrates that, conditional on the model, the weight of evidence in all the data overpowers the signal from the small sample for this drive-model, which suggests high rates infant mortality in the first quarter year.

For a clear example of this, consider drive-model 43. The data for this drive-model are heavily truncated; the observed units had run for quite some time before data were made available. In figure \ref{fig:ex-mod-43} we see that the lack of data for this drive-model at early ages results in a diffuse posterior centered around the what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_{\pi}, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$. Then, in the period where records exist for this drive model, at around $1\frac{1}{2}$ years, the posterior deviates from the global model, acheiving a compromise between it and the nonparametric estimate.

>>>>

\subsection{Results}
\subsubsection{Model Assessment}
To assess the goodness of fit for our final model (Model 4), we use a graphical posterior predictive check, wherein replicate data sets are generated from the posterior predictive distribution and compared with the observed data. This is a general approach recommended by \citet{gelman1996postpred} for models where no classical goodness-of-fit test is available. Because of the rich information gleaned from plotting the data, this method can be very effective at identifying limitations of the model.

Due to censoring, it is not entirely clear how to generate the replicate data sets because we have not modeled the censoring mechanism. Since we plan to compare nonparametric lifetime estimates, we chose to replicate as best as possible the pattern of censoring that was observed -- to generate the data conditional on the observed pattern of censoring. The problem in doing so is that the censoring mechanism itself is not fully observed. Our solution was to assume, for units whose failures were observed (and whose censoring time was not), that had they been right-censored, the censoring time would have been the latest censoring or failure time in the data for that drive model. We chose to generate 19 replicates for each drive model. This gives a 1/20 chance under the hypothesis of no difference that the most extreme data set would be the observed one and is not too many so as to produce an overly cluttered plot. Our method for generating replicate data sets, $y_{rep,g}^{(s)}$, $s = 1 \dots 19$, for each drive-model $g=1,\ldots,44$, is as follows:
\begin{enumerate}
\item Simulate $\tilde{y}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$.  Where each set of parameters, $\left( \pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)} \right) $, is a random draw from the full posterior distribution.
\item Repeat step 1 until $\tilde{y}_{rep,g,i} > t_{g,i}^L$.
\item If $c_{g,i}=0$, let $c_{rep,g,i}^(s)=\max \{y_{g,i}: i \ge 1\}$, otherwise let $c_{rep,g,i}^{(s)}=y_{g,i}$.\\
If $\tilde{y}_{rep,g,i}^{(s)}>m_{g,i}$, set $y_{rep,g,i}^{(s)}=c_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=1$, otherwise set $y_{rep,g,i}^{(s)}=\tilde{y}_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=0$.
\end{enumerate}

We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to reestimate the Kaplan-Meier nonparametric estimator. The resulting plots are displayed in figure \ref{fig:post-pred-KM}. For most drive-models, the estimates from the observed data appear to behave similarly to those from the replicated data. As discussed in \ref{subsec:ex1}, drive-model 14 has a discrepancy in the right tail, which we addressed there. Drive-models 4 and 34 appear to be exceptions as well. We will focus our attention on drive-model 4, but our diagnosis of 34 is similar, so the following comments extend to that drive-model as well.

@@

To better understand the reason for the discrepancy between the observed data for drive-model 4 and its posterior predictive distribution, we plot the raw data along with the truncation-adjusted nonparametric estimate of time to failure and the posterior estimate (see figure \ref{fig:ex-mod-4}). From these plots, it is clear that the large discrepancy between the posterior and the nonparametric estimate is driven by a small number of failures that occur in the first quarter year. Indeed, if we ignore records with observation periods that include the first quarter, the modified nonparametric estimate falls within the posterior 90\% pointwise credible band. Our trouble with fitting this drive-model is that while there is evidence of infant mortality in the relatively few units that were observed in the first quarter, the problem seems to go away much quicker than would be expected if there were a large number of units susceptible to the early failure mode as estimated by the global parameters, $\mu_1$ and $\sigma_1$. In any case, this example demonstrates that, conditional on the model, the weight of evidence in all the data overpowers the signal from the small sample for this drive-model, which suggests high rates infant mortality in the first quarter year.

Generally we consider this behavior a feature of hierarchical modeling. For another example of this, consider drive-model 43. The data for this drive-model are heavily truncated; the observed units had run for quite some time before data were made available. In figure \ref{fig:ex-mod-43} we see that the lack of data for this drive-model at early ages results in a diffuse posterior centered around the what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_{\pi}, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$. Then, in the period where records exist for this drive model, at around $1\frac{1}{2}$ years, the posterior deviates from the global model, acheiving a compromise between it and the nonparametric estimate.
